A Multiple-Set Overlapping-Domain Decomposed Monte Carlo Synthetic Acceleration Method for Linear Systems

### 2013-09-08 doc/sna2013/paper/slattery_evans_wilson.tex
fixing typos in sna

\begin{figure}[h!]
  \begin{center}
    \scalebox{0.7}{ \input{domain_to_domain.pdftex_t} }
  \end{center}
  \caption{Nearest neighbor history communication sequence.
    \textit{Each subdomain in the system has a set of nearest
      neighbors determined by the parallel adjacency graph of the
      input matrix. The subdomains are indexed by an integer.}}
  \label{fig:nearest_neighbor_comm}
\end{figure}
Given the mutual exclusivity of the transport algorithm, an additional
communcation sequence is required to efficiently stop the algorithm
after all histories have been computed. To stop transport, Brunner and
Brantley use an asynchronous binary communication tree as presented
for the same 9 subdomain example in
Figure~\ref{fig:binary_comm_tree}. In this communication pattern, each
process has one parent process (except for process 0) to which it will
non-blocking send the number of histories that terminated their
transport procedure by weight cutoff in the local
domain. Equivalently, each process has up to two child processes from
which it will receive their completed number of histories with a
non-blocking receive operation. Setting up a tree in this manner lets
the completed history tally be updated incrementally and funneled to
the root process in a mutually exclusive manner. Once the root process
has tallied a completed number of histories equal to the number of
histories in the source, transport is complete. Once this occurs, the
root process non-blocking sends the stop message to its children and
each child process non-blocking receives a stop message from the root
as shown in Figure~\ref{fig:binary_comm_tree}. The stop message is
then propagated up the tree in the same manner between parents and
children.
\begin{figure}[h!]
  \begin{center}
    \scalebox{0.6}{
      \input{binary_comm_tree.pdftex_t} }
  \end{center}
  \caption{Binary communication tree for coordinating the end of a
    parallel Neumann-Ulam solve. \textit{Each child process reports to
      its parents how many histories completed within its domain. When
      the root process history sum equals the total number of
      histories, it forwards the stop signal to its children which in
      turn forward the message up the tree. The subdomains are indexed
      by an integer.}}
  \label{fig:binary_comm_tree}
\end{figure}

@@

To implement the MSOD algorithm, we physically copy all data
structures required to implement the Neumann-Ulam algorithm to each of
the sets in the problem. Within each set, the subdomains detect their
neighboring subdomains and gather the number of adjacent states
specified for overlap. For each of these overlap states, the
probablities, weights, tallies, and other information required to
implement the Monte Carlo sequence are gathered. When transport
begins, within each individual set the Neumann-Ulam sequence is
executed with parallelism provided Brunner and Brantely's algorithm as
outlined in the previous section. Here, transport occurs both in the
local domain and the overlap region before histories are
communicated. Once transport in all sets is compelete, the tallies
from all overlapping states are gathered to their original owning
processors within each set and then all tallies are summed onto the
set of processors occupied by the linear problem to construct the
final tally vector.

>>>>

\begin{figure}[h!]
  \begin{center}
    \scalebox{0.7}{ \input{domain_to_domain.pdftex_t} }
  \end{center}
  \caption{Nearest neighbor history communication sequence.
    \textit{Each subdomain in the system has a set of nearest
      neighbors determined by the parallel adjacency graph of the
      input matrix. The subdomains are indexed by an integer.}}
  \label{fig:nearest_neighbor_comm}
\end{figure}
Given the mutual exclusivity of the transport algorithm, an additional
communication sequence is required to efficiently stop the algorithm
after all histories have been computed. To stop transport, Brunner and
Brantley use an asynchronous binary communication tree as presented
for the same 9 subdomain example in
Figure~\ref{fig:binary_comm_tree}. In this communication pattern, each
process has one parent process (except for process 0) to which it will
non-blocking send the number of histories that terminated their
transport procedure by weight cutoff in the local
domain. Equivalently, each process has up to two child processes from
which it will receive their completed number of histories with a
non-blocking receive operation. Setting up a tree in this manner lets
the completed history tally be updated incrementally and funneled to
the root process in a mutually exclusive manner. Once the root process
has tallied a completed number of histories equal to the number of
histories in the source, transport is complete. Once this occurs, the
root process non-blocking sends the stop message to its children and
each child process non-blocking receives a stop message from the root
as shown in Figure~\ref{fig:binary_comm_tree}. The stop message is
then propagated up the tree in the same manner between parents and
children.
\begin{figure}[h!]
  \begin{center}
    \scalebox{0.6}{
      \input{binary_comm_tree.pdftex_t} }
  \end{center}
  \caption{Binary communication tree for coordinating the end of a
    parallel Neumann-Ulam solve. \textit{Each child process reports to
      its parents how many histories completed within its domain. When
      the root process history sum equals the total number of
      histories, it forwards the stop signal to its children which in
      turn forward the message up the tree. The subdomains are indexed
      by an integer.}}
  \label{fig:binary_comm_tree}
\end{figure}

@@

To implement the MSOD algorithm, we physically copy all data
structures required to implement the Neumann-Ulam algorithm to each of
the sets in the problem. Within each set, the subdomains detect their
neighboring subdomains and gather the number of adjacent states
specified for overlap. For each of these overlap states, the
probabilities, weights, tallies, and other information required to
implement the Monte Carlo sequence are gathered. When transport
begins, within each individual set the Neumann-Ulam sequence is
executed with parallelism provided Brunner and Brantley's algorithm as
outlined in the previous section. Here, transport occurs both in the
local domain and the overlap region before histories are
communicated. Once transport in all sets is complete, the tallies
from all overlapping states are gathered to their original owning
processors within each set and then all tallies are summed onto the
set of processors occupied by the linear problem to construct the
final tally vector.
